Constructing Quotient Inductive-Inductive Types

### 2018-10-29 paper.tex
Expand related work on Cartmell, FoSSaCS QIIT

Cartmell \cite{gat} also uses a type-theoretic syntax for signatures
of a class of theories, called generalised algebraic theories. He mutually
defines the language for writing down signatures and the set of
well-formed signatures themselves, using notions of preterms,
well-formed rules and derived rules. We avoid this complication by
first fixing an internal syntax of signatures. However, we don't have
an analogous completeness result, which in our case would mean that
for each CwF$^\K_{\Eq}$ there is a corresponding signature. Also,
Cartmell works in ZFC, while we use Martin-Löf type theory as
metatheory.

@@

In \cite{altenkirch2018quotient} a
semantic analysis of QIITs using finite limits is given. However, this
notion does not enforce strict positivity.

>>>>

Cartmell \cite{gat} also used a type-theoretic syntax for a class of
theories, called generalised algebraic theories (GATs). GATs cover
roughly the same theories as this work does, with the difference that
GATs additionally allow equations between sorts. Besides this, there
are two main differences to the current work.

First, regarding the definitions of the signatures: GATs are given by
a nameful presyntax along with well-formedness relations, while our
signatures are intrinsically typed and given by structured
categories. This greatly simplifies formal development, makes
machine-checked formalisation feasible, and also clarifies the notion
of induction over signatures.

Second, the focus of the current work differs. Cartmell's main result
was the establishment of contextual categories as classifying
categories of GATs. In contrast, we do not discuss classifying
categories\footnote{This could be included in future work.}. Instead,
we present more constructions involving algebras; these include an
explicit initial algebra construction and the CwF model, which yields
a small internal type theory of algebras for each signature, and which
in particular allows us to exactly compute eliminators and
homomorphisms as types in the meta type theory.

\cite{altenkirch2018quotient} also concerns QIITs. There, signatures
are given by a scheme which builds up complete categories of algebras
from lists of functors. This notion of signature is more semantic than
our one, with more overhead in encoding signatures. It also does not
enforce strict positivity, hence it does not always support initial
algebras. Nonetheless, it is shown that well-behaved QIIT signatures
are covered by this scheme. Also, induction is shown to be equivalent
to initiality, but the notion of induction is given only up to
isomorphisms of algebras, in contrast to the current work, where it is
computed strictly.

@@
